### 2018-09-16 Chapters/Chapter1.tex
Transformation

>>>>

In this section we will show how a continuum one dimensional model can be obtained from the discrete Heisenberg Hamiltonian in order to describe the AFM order and its dynamics. We write the Heisenberg Hamiltonian as $\hat{H} = J \sum_{\langle i,j \rangle} \bs{S}_i\bs{S}_j - K \sum_i S_{iz}^2$ with $J>0$ and $K$ is the anisotropy energy. Let the lattice be a one dimensional chain with $2N$ sites, such lattice can be described by the repetition of $N$ unit cells consisting of two sites $\alpha$ and $\beta$ for the left and right site respectively. Then the ground state of this Hamiltonian is double degenerated and is obtained by aligning the spins in the $\alpha$ sites in the $\hat{z}$ direction and the $\beta$ spins in the $-\hat{z}$ direction or vice versa. Let us introduce the parameters:

\begin{align}
\bs{m}_i &= \frac{\bs{S}_{\alpha}^i + \bs{S}_{\beta}^i}{2S} \\
\bs{l}_i &= \frac{\bs{S}_{\alpha}^i - \bs{S}_{\beta}^i}{2S}
\end{align}

Where $S$ is the spin angular momentum in units of $\hbar$. Inverting these relations, introducing them into the Hamiltonian and neglecting edge terms we obtain:

\begin{align}
\hat{H} &= JS^2 \sum_i^{N-1} (\bs{m}_i-\bs{l}_i)(\bs{m}_i+\bs{l}_i+\bs{m}_{i+1}+\bs{l}_{i+1}) - KS^2\sum_i^N \left[(\bs{m}_{iz}+\bs{l}_{iz})^2 + (\bs{m}_{iz}-\bs{l}_{iz})^2 \right] \nonumber = \\
& 2JS^2\sum_i^{N} (\bs{m}_i^2-\bs{l}_i^2)+\frac{JS^2}{2}\sum_i^{N-1}\left[(\bs{l}_{i+1}-\bs{l}_i)^2-(\bs{m}_{i+1}-\bs{m}_i)^2 \right] \nonumber \\
& + JS^2\sum_i^{N-1} \left[ \bs{m}_i(\bs{l}_{i+1}-\bs{l}_i) - \bs{l}_i(\bs{m}_{i+1}-\bs{m}_i) \right] - 2KS^2\sum_i^N(m_{iz}^2+l_{iz}^2)
\end{align}

### 2018-09-16 Chapters/Chapter3.tex
Transformation

Which is the Hamiltonian \ref{Hubbard} without DMI interaction. We can introduce a time dependent electric field in this Hamiltonian by adding a phase to the hopping term $t_{ij}(t) = t_0e^{ieaA_{ij}(t)}$, where $A_{ij}$ is the projection of the vector potential into the vector from site $i$ to $j$. The electric field has the form $\bs{E}(t) = E_0 \sin (\omega t)$ (one-dimensional lattice). We can therefore take $A_{ij} = -\frac{E_0}{\omega}\cos(\omega t)(i-j)$.

The Hamiltonan with the new $t_{ij}(t)$ is periodic in time with period $T=\frac{2\pi}{\omega}$. According to Floquet's theorem there exist solutions that have the form $\ket{\psi(t)} = e^{-i \epsilon_\alpha t} \ket{\psi_\alpha (t)}$ where $\ket{\psi_\alpha(t)}$ has also period $T$, and so it can be expanded in Fourier series $\ket{\psi_\alpha(t)} = \sum_m e^{i\omega mt} \ket{\psi_{\alpha,m}}$ introducing this ansatz in the Scrodinger equation and projecting into the m'th mode we obtain:

\begin{equation}
(\epsilon_\alpha-\omega m) \ket{\psi_{\alpha, m}} = \sum_{m'} = H_{m'-m} \ket{\psi_{\alpha,m'}}
\end{equation}

Where $H_m = \frac{1}{T}\int_0^T dt e^{i\omega mt} H(t)$. The hopping term with the Peierls phase has Fourier components:

\begin{equation}
H_{m} = \frac{t_0}{T} \sum_{\langle i,j \rangle, \sigma} \int_0^T dt e^{i\omega mt} e^{-iea\frac{E_0}{\omega}\cos(\omega t)(i-j)}\hat{c}_{i\sigma}^\dagger \hat{c}_{j\sigma} = -t_0 \sum_{\langle i,j \rangle, \sigma} (-1)^mJ_m((i-j)ea\frac{E_0}{\omega})\hat{c}_{i\sigma}^\dagger \hat{c}_{j\sigma}
\end{equation}

Where $J_m(x)$ is the \textit{m}th Bessel function.

>>>>

Which is the Hamiltonian \ref{Hubbard} without DMI interaction. We can introduce a time dependent electric field in this Hamiltonian by adding a phase to the hopping term $t_{ij}(t) = t_0e^{ie\bs{R}_{ij}\bs{A}(t)}$, where $\bs{A}$ is the vector potential and $\bs{R}_{ij} = \bs{R}_i-\bs{R}_j$. The electric field has the form $\bs{E}(t) = E_0 \sin (\omega t)\bs{\hat{x}}$ (one-dimensional lattice). We can therefore take $\bs{A}(t) = -\frac{E_0}{\omega}\cos(\omega t)\bs{\hat{x}} = -\bs{A}\cos(\omega t)$. Therefore the time dependent Hamiltonian is:

\begin{equation}
\hat{H}(t) = -\sum_{\langle i,j \rangle, \sigma} t_{ij}(t)\hat{c}_{i \sigma}^\dagger \hat{c}_{j \sigma} + U \sum_{i=1}^M \hat{n}_{i\uparrow}\hat{n}_{i\downarrow} = -\hat{T}(t) + U\hat{D}
\end{equation}

Where $\hat{D}$ is the double occupancy number operator and $\hat{T}(t) = \sum_{\langle i,j \rangle, \sigma} t_{ij}(t)\hat{c}_{i \sigma}^\dagger \hat{c}_{j \sigma}$ is the hopping operator. Since the time dependence in periodic we can expand $t_{ij}(t) = \sum_m t_{ij,m} e^{im \omega t}$, where

\begin{equation}
t_{ij,m} = t_0(-1)^mJ_m(\alpha_{ij})
\end{equation}

Where $\alpha_{ij} = e\bs{R}_{ij}\bs{A}$ and $J_m(x)$ is the \textit{m}th Bessel function, and correspondingly $\hat{T}(t) = \sum_m \hat{T}_m e^{im \omega t}$.

In order to derive the form of the effective Hamiltonian in the half filling system with $\frac{t}{U} << 1$ let us introduce a time dependent unitary transformation $\hat{U}(t) = e^{-i\hat{S}(t)}$. The transformed Hamiltonian is:

\begin{equation}
\hat{H}'(t) = \hat{U}^\dagger (t) \hat{H}(t) \hat{U}(t) - i\hat{U}^\dagger(t) d_t \hat{U}(t)
\end{equation}

We perform the unitary transformation perturbatively in the hopping parameter $t_0$. We expand $\hat{S}(t) = \sum_\nu \hat{S}^{(\nu)}(t)$ and $\hat{H}'(t) = \sum_\nu \hat{H}'^{(\nu)}(t)$. In order for the new Hamiltonian to be periodic we impose the unitary transformation to have the periodicity of the Hamiltonian, so that we can expand $\hat{S}^{(\nu)}(t) = \sum_m e^{im\omega t}\hat{S}^{(\nu)}_m$. Additionally we can impose the transformed Hamiltonian to be block diagonal in the doublon number $d$. With these conditions the unitary transformation can be uniquely determined if we impose that $\hat{S}(t)$ does not contain block-diagonal terms, i.e. we can write:

\begin{equation}
\hat{S}^{(\nu)}(t) = \sum_{d \neq 0} \sum_m \hat{S}^{(\nu)}_{d,m} e^{im\omega t}
\end{equation}

where $\hat{S}^{(\nu)}_{d,m}$ changes the double occupancy number by $d$. We also decompose the hopping operator as:

\begin{equation}
\hat{T}(t) = \sum_m (\hat{T}_{-1,m}+\hat{T}_{0,m}+\hat{T}_{1,m})e^{im\omega t}
\end{equation}

Where $\hat{T}_d(t)$ changes the double occupancy number by $d$, for example $\hat{T}_1(t) = \hat{P}_1 \hat{T}(t) \hat{P}_0$. We can write the transformed Hamiltonian as:

\begin{equation}
\label{Perturbation}
\hat{H}'(t) = \hat{H}(t) + \sum_{n=1}^\infty \frac{1}{n!} \left[i\hat{S}(t), \dots, \left[ i\hat{S}(t), \hat{H}(t) \right]\dots \right] - d_t \hat{S}(t)
\end{equation}

To zeroth order we have $\hat{H}'^{(0)}= U\hat{D}$ and $i\hat{S}^{(0)}(t)=0$. In first order we obtain:

\begin{equation}
\hat{H}'^{(1)}(t)=-\hat{T}(t)-\sum_{d\neq 0}\sum_m (Ud+m\omega) i\hat{S}^{(1)}_{dm} e^{im\omega t}
\end{equation}

Therefore:

\begin{align*}
i\hat{S}^{(1)}_d(t) &= -\sum_m \frac{\hat{T}_{d,m}}{Ud+m\omega}e^{im\omega t} \\
\hat{H}'^{(1)}(t) &= -\hat{T}_0(t)
\end{align*}

To second order we find:

\begin{equation}
\hat{H}'^{(2)}(t) = \left[i\hat{S}^{(2)}(t), U \hat{D} \right] - \left[ i\hat{S}^{(1)}(t), \hat{T}(t) \right] - d_t\hat{S}^{(2)}(t)
\end{equation}

Where the middle term is:

\begin{align*}
\left[ i\hat{S}^{(1)}(t), \hat{T}(t) \right] &= -\left[\sum_m \left( \frac{\hat{T}_{1m}}{U+m\omega} - \frac{\hat{T}_{-1m}}{U-m\omega} \right)e^{im \omega t}, \sum_n \left( \hat{T}_{-1n} +\hat{T}_{0n} + \hat{T}_{1n} \right) e^{in\omega t} \right] \\
&= -\sum_{mn} \left\{ \frac{\left[\hat{T}_{1m}, \hat{T}_{0n} \right]}{U+m\omega} + \frac{\left[\hat{T}_{1m}, \hat{T}_{-1n} \right]}{U+m\omega} - \frac{\left[\hat{T}_{-1m}, \hat{T}_{1n} \right]}{U-m\omega} - \frac{\left[\hat{T}_{-1m}, \hat{T}_{0n} \right]}{U-m\omega} \right\} e^{i(m+n)\omega t} \\
&= -\sum_{mn} \left\{ \frac{\left[\hat{T}_{1n}, \hat{T}_{0(m-n)} \right]}{U+n\omega} + \frac{\left[\hat{T}_{1n}, \hat{T}_{-1(m-n)} \right]}{U+n\omega} - \frac{\left[\hat{T}_{-1n}, \hat{T}_{1(m-n)} \right]}{U-n\omega} - \frac{\left[\hat{T}_{-1n}, \hat{T}_{0(m-n)} \right]}{U-n\omega} \right\} e^{im\omega t}
\end{align*}

We obtain:

\begin{align*}
i\hat{S}^{(2)}_{dm} &= \sum_n \frac{\left[ \hat{T}_{dn}, \hat{T}_{0(m-n)} \right]}{(Ud+n\omega)(Ud+m\omega)} \\
\hat{H}'^{(2)}(t) &= \sum_{mn} \left( \frac{\left[\hat{T}_{1n}, \hat{T}_{-1(m-n)} \right]}{U+n\omega} - \frac{\left[\hat{T}_{-1n}, \hat{T}_{1(m-n)} \right]}{U-n\omega} \right) e^{im\omega t}
\end{align*}

Now the effective Hamiltonian acts on the subspace of $\hat{P}_0$, therefore we are only interested on $\hat{P}_0 \hat{H}'^{(2)}(t) \hat{P}_0$ (notice that $\hat{P}_0 \hat{H}'^{(1)}(t) \hat{P}_0 = 0$). In this subspace we can write:

\begin{align*}
\hat{T}_{1n} \hat{T}_{-1(m-n)} &= 0 \\
\hat{T}_{-1(m-n)} \hat{T}_{1n} &= \sum_{\langle i,j \rangle, \sigma_1, \sigma_2} t_{ji,(m-n)} t_{ij,n} \hat{c}_{j \sigma_2}^\dagger \hat{c}_{i \sigma_2} \hat{c}_{i \sigma_1}^\dagger \hat{c}_{j \sigma_1}
\end{align*}

So that,

\begin{equation}
\hat{H}_{eff}(t) = \hat{P}_0\hat{H}'^{(2)}(t)\hat{P}_0 = - \sum_{mn} \sum_{\langle i,j \rangle, \sigma_1, \sigma_2}\hat{c}_{j \sigma_2}^\dagger \hat{c}_{i \sigma_2} \hat{c}_{i \sigma_1}^\dagger \hat{c}_{j \sigma_1} \left( \frac{t_{ji,(m-n)} t_{ij,n}}{U+n\omega} + \frac{t_{ji,n} t_{ij,(m-n)}}{U-n\omega} \right) e^{im\omega t}
\end{equation}

Comparing with \ref{PertNoEMF} we can see that this will lead to a spin Hamiltonian

\begin{equation}
\hat{H}_{eff}(t) = \sum_{\langle i,j \rangle} J_{ij}(t) \bs{S}_i \bs{S}_j
\end{equation}

With

\begin{align}
J_{ij}(t) &= \sum_{mn} \left( \frac{t_{ji,(m-n)} t_{ij,n}}{U+n\omega} + \frac{t_{ji,n} t_{ij,(m-n)}}{U-n\omega} \right) e^{im\omega t} \nonumber \\
&= \sum_{mn} (-1)^m t_0^2 \left( \frac{J_{m-n}(\alpha_{ji})J_n(\alpha_{ij})}{U+n\omega} +\frac{J_n(\alpha_{ji})J_{(m-n)}(\alpha_{ij})}{U-n\omega} \right)
\end{align}

In the case $\omega >> U$ we can limit to $m=0$ (time independent part) and $n=0$ or $n=\pm1$ (small denominators), additionally, if the electric field is small we can use $J_n(x) \aprox x^n$. With these approximations:

\begin{equation}
J_{ij} = 2 t_0^2 \alpha_{ij}^2 \left( \frac{1}{U-\omega} - \frac{1}{U} + \frac{1}{U+\omega} \right)
\end{equation}
